Sequential and Parallel Solution-Biased Search for Subgraph Algorithms

### 2019-06-07 talk/talk.tex
Tweaking

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.4cm)}] at (current page.north east) {
            \begin{minipage}{0.20\paperwidth}
                \begin{center}
                \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
                \includegraphics[keepaspectratio=true,scale=0.8]{UStA_Logo_SP.eps}
                \end{center}
            \end{minipage}
        };

@@

\begin{frame}{Subgraph Isomorphism}
    \centering
    \begin{tikzpicture}
        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

@@

        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw [ultra thick] (N1) -- (N3);
        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw [ultra thick] (N3) -- (N5);
        \draw [ultra thick] (N3) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw [ultra thick] (N5) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Subgraph Finding, as a Constraint Program}
    \begin{itemize}
        \item A variable for each pattern vertex. The domains are all of the target vertices.
        \item At least two sets of constraints:
            \begin{itemize}
                \item Adjacent pairs of vertices must be mapped to adjacent pairs of vertices.
                \item All different.
            \end{itemize}
        \item Then we get clever:
            \begin{itemize}
                \item Extra constraints about degrees, paths, \ldots
                \item Very good variable- and value-ordering heuristics
            \end{itemize}
    \end{itemize}
\end{frame}

@@

    \begin{itemize}
        \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
            locally injective, labels, side constraints, directed, \ldots).
        \item Also special algorithms for clique.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Sanity Check}
    \begin{center}
        \includegraphics<1>{gen-graph-value-ordering.pdf}%
        \includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
    \end{center}
\end{frame}

@@

\begin{frame}{A Slightly Random Value-Ordering Heuristic}
    \only<1>{
    \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
    \begin{itemize}
        \item Equally likely to pick between two vertices of degree $d$.
        \item Twice as likely to select a vertex of degree $d$ than a vertex of degree $d - 1$.
        \item In the paper: \textcolor{uofgcobalt}{solution density} and expected distribution of solutions.
    \end{itemize}}\only<2>{
    \begin{center}
        \includegraphics<2>{gen-graph-bias-scatter.pdf}%
    \end{center}}
\end{frame}

@@

\begin{frame}{Broader Perspective}
    \begin{itemize}
        \item Also good for maximum common subgraph algorithms.
        \item Does this work for CP in general?
        \item Can we finally kill off vanilla sequential backtracking search?
    \end{itemize}
\end{frame}

>>>>

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.4cm)}] at (current page.north east) {
            \begin{minipage}{0.20\paperwidth}
                \begin{center}
                \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
                \includegraphics[keepaspectratio=true,scale=0.07]{UStA_Logo_SP.png}
                \end{center}
            \end{minipage}
        };

@@

\begin{frame}{Subgraph Isomorphism}
    \centering
    \begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

@@

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [ultra thick] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Subgraph Finding, as a Constraint Program}
    \begin{itemize}
        \item A variable for each pattern vertex. The domains are all of the target vertices.
        \item At least two sets of constraints:
            \begin{itemize}
                \item Adjacent pairs of vertices must be mapped to adjacent pairs of vertices.
                \item All different.
            \end{itemize}
        \item Then we get clever:
            \begin{itemize}
                \item Extra constraints about degrees, paths, \ldots
                \item Better data structures and propagation queues.
                \item Very good variable- and value-ordering heuristics.
            \end{itemize}
    \end{itemize}
\end{frame}

@@

    \begin{itemize}
        \item A CP style solver specifically for subgraph algorithms.
        \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
            locally injective, labels, side constraints, directed, \ldots).
        \item Also special algorithms for clique.
    \end{itemize}
\end{frame}

\begin{frame}{Benchmark Instances}
    \begin{itemize}
        \item 14,621 instances from Christine Solnon's collection:
            \begin{itemize}
                \item Randomly generated with different models.
                \item Real-world graphs.
                \item Computer vision problems.
                \item Biochemistry problems.
                \item Phase transition instances.
            \end{itemize}
        \item At least\ldots
            \begin{itemize}
                \item $\ge$ 2,110 satisfiable.
                \item $\ge$ 12,322 unsatisfiable.
            \end{itemize}
        \item A lot of them are very easy for good algorithms.
    \end{itemize}
\end{frame}

\begin{frame}{Hardware}
    \begin{itemize}
        \item HPC, optimised for throughput not reproducibility.\footnote{This work used the Cirrus
            UK National Tier-2 HPC Service at EPCC (http://www.cirrus.ac.uk) funded by the
            University of Edinburgh and EPSRC (EP/P020267/1)}
        \item Dual Intel Xeon E5-2695 v4 CPUs, $2 \times 18$ cores
        \item 256GBytes RAM
        \item GCC 7.2.0
        \item C++ native threads, SGI MPT MPI
    \end{itemize}
\end{frame}

@@

\begin{frame}{Sanity Check}\includegraphics<1>{gen-graph-value-ordering.pdf}\includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
\end{frame}

@@

\begin{frame}{A Slightly Random Value-Ordering Heuristic}
    \only<1>{
    \begin{itemize}
        \item For a fixed domain $D_v$, pick a vertex $v'$ from a domain $D_v$ with probability
    \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
        \item Equally likely to pick between two vertices of degree $d$.
        \item Twice as likely to select a vertex of degree $d$ than a vertex of degree $d - 1$.
        \item In the paper: \textcolor{uofgcobalt}{solution density} and expected distribution of solutions.
    \end{itemize}}\only<2>{
    \begin{center}
        \includegraphics<2>{gen-graph-bias-scatter.pdf}%
    \end{center}}
\end{frame}

@@

\begin{frame}{Broader Perspective}
    \begin{itemize}
        \item Can scale to even more hosts with less strict synchronisation.
        \item Also good for maximum common subgraph algorithms.
        \item Does this work for CP in general?
        \item Can we finally kill off vanilla sequential backtracking search?
    \end{itemize}
\end{frame}
